---
title: "Practical Machine Learning Course Project (Week 4 Assignment)"
author: "Wipada Treebumrungsak"
date: "14 March 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data

The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

## Environment Preparation
First, to upload R libraries that are necessary for week4 assignment

```{r environment}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(randomForest)

```

## Data Loading and Cleaning

- Download the data files from the Internet
- Load data into two data frames. 
- So, we have a training dataset and a 20 observations testing dataset
- Remove NA columns and remove features that are not in test dataset

```{r dataprep}
UrlTrain <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
UrlTest  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

# download the datasets
dt_training <- read.csv(url(UrlTrain))
dt_testing  <- read.csv(url(UrlTest))

#Cleaning the data
features <- names(dt_testing[,colSums(is.na(dt_testing)) == 0])[8:59]

# Only use features used in testing cases.
dt_training <- dt_training[,c(features,"classe")]
dt_testing <- dt_testing[,c(features)]

dim(dt_training); dim(dt_testing);

```

## Data Partition

split data into a training data set (60% of the total cases) and a testing data set (40% of the total cases)

```{r dataPartition}
set.seed(12345)

inTrain <- createDataPartition(dt_training$classe, p=0.6, list=FALSE)
training <- dt_training[inTrain,]
testing <- dt_training[-inTrain,]

dim(training)
```

```{r dataPartition2}
dim(testing)
```


## Model A: Build Decision Tree Model
Using Decision Tree, the accuracy might not be high. 
(Accuracy 80% is acceptable.)


```{r decisiontree}
modFitDT <- rpart(classe ~ ., data = training, method="class")
rpart.plot(modFitDT)
```


## Model A: Predict with Decision Tree Model

```{r predict_DT}
set.seed(12345)

prediction <- predict(modFitDT, testing, type = "class")
confusionMatrix(prediction, testing$classe)
```

## Model B: Build Random Forest Model

Using random forest, the out of sample error should be small. The error will be estimated using the 40% testing sample. We should expect an error estimate of < 3%.


```{r build_RF}

set.seed(12345)
modFitRF <- randomForest(classe ~ ., data = training, ntree = 1000)

```

## Model B: Predict with Random Forest Model

```{r predict_RF}

prediction <- predict(modFitRF, testing, type = "class")
confusionMatrix(prediction, testing$classe)

```

## Applying the Selected Model to the Test Data (pml-testing.csv)
The accuracy of the 2 predictive modeling methods above are:

Decision Tree : 0.7267
Random Forest : 0.9939

So, the Random Forest model will be applied to predict the 20 quiz results (testing dataset) as shown below.


```{r predictTest_RF}

predictionRF <- predict(modFitRF, dt_testing, type = "class")
predictionRF

```

